---
title: "GSEA Simulation report"
description: "We compare the performance of logistic-IBSS to linear SuSiE, RSS with logistic regression summary stats, and others on a set of simulations meant to emulate realistic case-control data in GWAS. We aim to identify regimes where the logistic regression model is necessary."
output: 
    html_document:
        toc: true
        toc_float: true
        toc_depth: 3
        code_folding: hide 
---

```{r setup}
#project_dir <- snakemake@config$project_dir
project_dir <-  '/project2/mstephens/ktayeb/logistic-susie-snakemake'
knitr::opts_knit$set(root.dir = project_dir)
```

```{r setup2}
library(dplyr)
library(dplyr)
library(reticulate)
library(tidyr)
library(tictoc)
library(ggplot2)

source('workflow/scripts/plotting/plot_functions.R')

# httpgd::hgd()

load_tbl <- function(path){
    print(path)
    dplyr::as_tibble(readRDS(path))
}
pip_summary_files <- Sys.glob('results/gsea/*/*_pip_summary.rds')
res_pip <- purrr::map_dfr(pip_summary_files, ~load_tbl(.x))

cs_summary_files <- Sys.glob('results/gsea/*/*_cs_summary.rds')
res_cs <- purrr::map_dfr(cs_summary_files, ~dplyr::as_tibble(readRDS(.x)))

manifest <- readr::read_delim('config/gsea_sim/gsea_sim_manifest.tsv', delim='\t')
```

```{r plot-funs}
threshold_var <- function(df, var, val, comp){
#   var <- as_string(ensym(var))
  filter(df, comp(.data[[var]], !!val))
}

cs_filtered_coverage <- function(res_cs, var, val, comp){
    Q_tbl <- res_cs %>%
        select(c(method, sim_id, q)) %>%
        unique() %>%
        group_by(method) %>%
        summarize(Q = sum(q))

    res_cs %>%
        left_join(Q_tbl) %>%
        threshold_var(var, val, comp) %>%
        mutate(val = val) %>%
        group_by(method) %>%
        summarize(coverage = mean(covered), n_found2 = list(table(n_found)), mean_size = mean(size), val = val[1], Q=Q[1], n_found = sum(n_found)) %>%
        unnest_wider(n_found2) 
}

make_plot <- function(res_cs, var, vals, op){
    f <- function(val){
        res_cs %>%
            cs_filtered_coverage(var, val, op)
    }
    filtered_coverage <- purrr::map_dfr(vals, f)

    a <- filtered_coverage %>%
        ggplot(aes(x = val, y = coverage, color = method)) +
        geom_path() + 
        geom_hline(yintercept = 0.95) +
        xlab(var)

    b <- filtered_coverage %>%
        mutate(power = n_found/Q) %>%
        mutate(power = `1`/Q) %>%
        ggplot(aes(x = val, y = power, color = method)) +
        geom_path() +
        xlab(var)

    cowplot::plot_grid(a, b)
}

```
## SER simulations

### Simulation settings

```{r}
# show unique simulation settings for SER simulations
manifest %>%
    filter(q == 1) %>%
    select(c(X, beta0, beta)) %>%
    unique() %>%
    rmarkdown::paged_table()
```

### Credible sets

#### CS size distribution 
```{r}
# histogram of CSs, and coverage over all simulations
res_cs %>%
    filter(q == 1) %>%
    ggplot(aes(x = size)) + 
    geom_histogram() +
    facet_wrap(vars(method))
```

#### CS coverage

```{r}
res_cs %>%
    filter(q == 1) %>%
    group_by(method) %>%
    summarize(coverage = mean(covered), n_found = list(table(n_found)), mean_size = mean(size)) %>%
    unnest_wider(n_found) %>%
    knitr::kable()
```

The Laplace ABF gives lower coverage marginally over all CSs and all simulation scenarios.
This is resolved by filtering out large CSs, which are largely uninformative.

```{r}
res_cs %>%
    filter(q == 1) %>%
    filter(size < 200) %>%
    group_by(method) %>%
    summarize(coverage = mean(covered), n_found = list(table(n_found)), mean_size = mean(size)) %>%
    unnest_wider(n_found) %>%
    knitr::kable()
```

#### Conditional coverage

```{r}
# filter on purity
vals <- c(seq(0.1, 0.9, by = 0.1), seq(0.91, .99, by = 0.01))
res_cs %>% make_plot('purity', vals, `>`)

# filter on cs size
vals <- seq(5, 100, by=5)
res_cs %>% make_plot('size', vals, `<`)

# filter on the SER-level BFs
# note multiple CSs may capture the same effect causing "power" > 1
vals <- seq(0, 100, by=5)
res_cs %>% make_plot('lbf_ser', vals, `>`)
```

### Power vs FDP (PIP ordering)

```{r}
# fdp vs power
fdp_plot <- res_pip %>%
    unnest_longer(c(pip, causal)) %>%
    ggplot(aes(pip, causal, color=method)) + 
    stat_fdp_power(geom='path', max_fdp=1.0) +
    labs(x = 'FDP', y = 'Power')
fdp_plot
```

### PIP Calibration 

PIP calibration looks a little messy-- may need more simulations to get a clear picture.
Note that for the logistic SERs the only source of error here is the approximation error of using the Laplace ABF vs Wakefield ABF.
TODO: add comparison to exact computation of the BF via quadrature.

```{r}
pip_calibration_plot <- res_pip %>%
    unnest_longer(c(pip, causal)) %>%
    ggplot(aes(pip, causal)) +
    stat_pip_calibration() +
    geom_abline(intercept = 0, slope = 1, color='red') +
    theme_bw() +
    labs(x='PIP', y='Frequency') + 
    facet_wrap(vars(method))
pip_calibration_plot
```
